# Performance Overview

## Comparing Peak Performance of Language Implementations

```{r prepare-data, echo=FALSE, message=TRUE, warning=TRUE, dev='svg'}
## Basic Setup
# load libraries, the data, and prepare it
if (Sys.getenv("RSTUDIO") == "1") { setwd("/Users/smarr/Projects/PostDoc/FASTXX/are-we-fast-yet/report") }

source("scripts/libraries.R", chdir=TRUE)
opts_chunk$set(dev = 'svg')

data <- load_data_file("data/benchmark.data")
data <- subset(data, select = c(Value, Unit, Benchmark, VM, Iteration))
#data <- prepare_vm_names(data)

# summary(data)
# levels(data$VM)

jit <- droplevels(subset(data, Iteration >= 700 & Iteration <= 999))
mri <- droplevels(subset(data, VM == "MRI22"))
ruby_indy <- droplevels(subset(data, Iteration >= 90 & Iteration <= 170 & VM %in% c("JRubyC2", "JRubyGraal", "JRubyJ8")))
crystal <- droplevels(subset(data, Iteration >= 200 & Iteration <= 400))

peak <- rbind(jit, mri, ruby_indy, crystal)

norm <- ddply(peak, ~ Benchmark, transform,
              RuntimeRatio = Value / mean(Value[VM == "GraalEnterprise"]))
stats <- ddply(norm, ~ VM + Benchmark, summarise,
               Time.ms = mean(Value),
               sd      = sd(Value),
               RuntimeFactor = geometric.mean(RuntimeRatio),
               RR.sd         = sd(RuntimeRatio))
stats <- droplevels(subset(stats, Benchmark != "DeltaBlueV8"))

norm_j8 <- ddply(peak, ~ Benchmark, transform,
              RuntimeRatio = Value / mean(Value[VM == "Java8U66"]))
stats_j8 <- ddply(norm_j8, ~ VM + Benchmark, summarise,
               Time.ms = mean(Value),
               sd      = sd(Value),
               RuntimeFactor = geometric.mean(RuntimeRatio),
               RR.sd         = sd(RuntimeRatio))
stats_j8 <- droplevels(subset(stats_j8, Benchmark != "DeltaBlueV8"))


plot_benchmarks_speedup_for_vms <- function(stats, vms) {
  vm_stats <- droplevels(subset(stats, VM %in% vms))
  
  for (b in levels(vm_stats$Benchmark)) {
    data_b <- droplevels(subset(vm_stats, Benchmark == b))
    
    p <- ggplot(data_b, aes(x = VM, y = RuntimeFactor, fill = VM)) +
      geom_bar(stat = "identity") +
      geom_errorbar(aes(ymax = RuntimeFactor + RR.sd, ymin = RuntimeFactor - RR.sd), width=0.25) +
      coord_flip() + theme_bw() + # scale_fill_manual(values=col) +
      theme(legend.position="none") + ggtitle(b)
    tryCatch({print(p)})
  }
}

plot_benchmarks_speedup_for_vms_faceted <- function(stats, vms) {
  vm_stats <- droplevels(subset(stats, VM %in% vms))
    
  p <- ggplot(vm_stats, aes(x = VM, y = RuntimeFactor, fill = VM)) +
      geom_bar(stat = "identity") +
      geom_errorbar(aes(ymax = RuntimeFactor + RR.sd, ymin = RuntimeFactor - RR.sd), width=0.25) +
      facet_wrap(~ Benchmark, ncol = 1, scales="free_y") +
       theme_bw() + theme_simple() + # scale_fill_manual(values=col) + coord_flip() +
      theme(legend.position="none", axis.text.x=element_text(angle=90, hjust = 1, vjust = 0.5))
  print(p)
}

overview_box_plot <- function(stats, vms, prepare_data = NULL) {
  cat(vms)
  vm_stats <- droplevels(subset(stats, VM %in% vms))
  
  if (!is.null(prepare_data)) {
    vm_stats <- prepare_data(vm_stats)
  }
  
  plot <- ggplot(vm_stats, aes(x=VM, y=RuntimeFactor, fill = VM)) +
    #geom_hline(aes(yintercept=1), colour="#cccccc", linetype="dashed") +
    #geom_hline(aes(yintercept=5), colour="#cccccc", linetype="dashed") +
    geom_boxplot() + #fill=get_color(5, 7)
    theme_bw() + theme_simple() + theme(axis.text.x = element_text(angle= 90, vjust=0.5, hjust=1), legend.position="none") +
    #scale_y_log10(breaks=c(1,2,3,10,20,30,50,100,200,300,500,1000)) + #limit=c(0,30), breaks=seq(0,100,5), expand = c(0,0)
    ggtitle("Runtime Factor, normalized to Java 1.8.0_66 (lower is better)") + coord_flip() + xlab("")
  
    #labeller = label_parsed
  print(plot)
}
```

The following plots show the runtime factor over GraalEnterprise (Java benchmark version)

 - GraalEnterprise: top-tier JIT = Enterprise Graal
 - GraalC2: top-tier JIT = C2
 - GraalBasic: top-tier JIT = Basic Graal
 - GraalJS: on Enterprise Graal
 - JRubyTruffle: on Basic Graal
 - JRubyTruffleEnterprise: on Enterprise Graal
 - JRubyTruffleMaster07: JRuby master branch on Graal 0.7 release
 - Node: node.js 5.2
 - SOMns: on Basic Graal

The benchmarks are implemented as identical as possible, trying to exercise a
'core' language common to the benchmarked system.
More details on the methodology and all benchmarks are available in the
[Are We Fast Yet?](https://github.com/smarr/are-we-fast-yet#are-we-fast-yet) repository.

### Are Truffle Languages Fast Yet? (Compared to Java 1.8.0_66)
<a id="all"></a>

```{r truffle-langs-overview, echo=FALSE, fig.keep='all', fig.width=8, fig.height=3}
overview_box_plot(stats_j8, c("GraalEnterprise", "GraalJS", "JRubyTruffle", "JRubyTruffleEnterprise", "JRubyTruffleMaster07", "SOMns", "SOMns-Enterprise"))


overview_box_plot(stats_j8, c("Java8U66", "GraalJS", "JRubyTruffleEnterprise", "SOMns"))
overview_box_plot(stats_j8, c("Java8U66", "GraalJS", "JRubyTruffle", "SOMns"))

overview_box_plot(stats_j8, c("Java8U66", "JRubyJ8", "Node", "MRI22"))
```

<a id="candidate-plot"></a>
```{r candidate-plot, echo=FALSE, fig.keep='all', fig.width=8, fig.height=3}

prep <- function (d) {
  d <- prepare_vm_names(d)
  d$VM <- reorder(d$VM, X=-d$RuntimeFactor)
  d
}
overview_box_plot(stats_j8, c("Java8U66", "JRubyJ8", "Node", "MRI22", "SOMns", "JRubyTruffleEnterprise"), prep)
```

```{r truffle-lang-table, results='asis', echo=FALSE}
vm_stats_j8 <- ddply(stats_j8, ~ VM, summarise,
                     geomean = geometric.mean(RuntimeFactor),
                     sd      = sd(RuntimeFactor),
                     min     = min(RuntimeFactor),
                     max     = max(RuntimeFactor),
                     median  = median(RuntimeFactor))
vm_stats_j8$VM <- reorder(vm_stats_j8$VM, X=vm_stats_j8$geomean)
#vm_stats_j8_o <- arrange(vm_stats_j8, Runtime.Factor)


t <- tabular(Justify("l")*Heading()*VM ~ 
             Heading('Runtime Factor over Java8U66')*Justify("r")*Format(sprintf("%.2f"))*((geomean + sd + min + max + median)*Heading()*identity), data=vm_stats_j8)
table_options(justification="c ")
html(t)
```

```{r truffle-langs, echo=FALSE, fig.keep='all', fig.width=4, fig.height=16}
plot_benchmarks_speedup_for_vms_faceted(stats, c("GraalEnterprise", "GraalJS", "JRubyTruffleEnterprise", "SOMns"))

#plot_benchmarks_speedup_for_vms_faceted(stats, c("GraalEnterprise", "GraalJS", "JRubyTruffle", "JRubyTruffleEnterprise", "JRubyTruffleMaster07", "SOMns", "SOMns-Enterprise"))
```

### Is Graal Fast Yet? (Baseline Graal Enterprise)
<a id="graal"></a>

```{r graal-overview, echo=FALSE, fig.keep='all', fig.width=8, fig.height=3}
overview_box_plot(stats, c("GraalBasic", "GraalC2", "GraalEnterprise", "Java8U66"))
```

```{r graal, echo=FALSE, fig.keep='all', fig.width=3, fig.height=16}
# vms <- c("GraalBasic", "GraalC2", "GraalEnterprise")
plot_benchmarks_speedup_for_vms_faceted(stats, c("GraalBasic", "GraalC2", "GraalEnterprise", "Java8U66"))
```

### Is Graal Fast Yet? (JRuby-Indy on C2 or Graal)

```{r graal-jruby-indy-overview, echo=FALSE, fig.keep='all', fig.width=8, fig.height=3}
overview_box_plot(stats, c("JRubyC2", "JRubyGraal", "JRubyJ8"))
```


```{r graal-jruby-indy, echo=FALSE, fig.keep='all', fig.width=2, fig.height=16}
plot_benchmarks_speedup_for_vms_faceted(stats, c("JRubyC2", "JRubyGraal", "JRubyJ8"))
```

### Is Graal.JS Fast Yet?
<a id="js"></a>

```{r graaljs-overview, echo=FALSE, fig.keep='all', fig.width=8, fig.height=3}
overview_box_plot(stats, c("GraalEnterprise", "GraalJS", "Node"))
```

```{r graaljs, echo=FALSE, fig.keep='all', fig.width=3, fig.height=16}
plot_benchmarks_speedup_for_vms_faceted(stats, c("GraalEnterprise", "GraalJS", "Node"))
```

### Is JRuby+Truffle Fast Yet?
!!!*TODO*: fix chart titles, normally the base line is GraalEnterprise except for some charts at the very top
<a id="ruby"></a>

```{r jruby-truffle-overview, echo=FALSE, fig.keep='all', fig.width=8, fig.height=3}
overview_box_plot(stats, c("GraalEnterprise", "JRubyTruffle", "JRubyTruffleEnterprise", "JRubyTruffleMaster07"))
```

```{r jruby-truffle, echo=FALSE, fig.keep='all', fig.width=4, fig.height=16}
plot_benchmarks_speedup_for_vms_faceted(stats, c("GraalEnterprise", "JRubyTruffle", "JRubyTruffleEnterprise", "JRubyTruffleMaster07"))
```

Including JRuby-Indy (w/o Truffle) and MRI
```{r jruby-truffle-overview-all, echo=FALSE, fig.keep='all', fig.width=8, fig.height=3}
overview_box_plot(stats,  c("GraalEnterprise", "JRubyC2", "JRubyGraal", "JRubyJ8", "JRubyTruffle", "JRubyTruffleEnterprise", "JRubyTruffleMaster07", "MRI22"))
```

```{r jruby-truffle-all, echo=FALSE, fig.keep='all', fig.width=5, fig.height=16}
plot_benchmarks_speedup_for_vms_faceted(stats, c("GraalEnterprise", "JRubyC2", "JRubyGraal", "JRubyJ8", "JRubyTruffle", "JRubyTruffleEnterprise", "JRubyTruffleMaster07", "MRI22"))
```


# Other Plots

## Absolute Runtime

```{r fast-vms-ms, echo=FALSE, fig.keep='all', fig.width=10, fig.height=6}
fast_stats <- droplevels(subset(stats, VM != "MRI22" & VM != "JRubyGraal" & VM != "JRubyC2" & VM != "JRubyJ8"))
for (b in levels(fast_stats$Benchmark)) {
  data_b <- droplevels(subset(fast_stats, Benchmark == b))
  
  p <- ggplot(data_b, aes(x = VM, y = Time.ms, fill = VM)) +
    geom_bar(stat = "identity") +
    geom_errorbar(aes(ymax = Time.ms + sd, ymin = Time.ms - sd), width=0.25) +
    coord_flip() + theme_bw() + # scale_fill_manual(values=col) +
    theme(legend.position="none") + ggtitle(b)
  print(p)
}
```

## Including Slow Implementations

```{r all-vms, echo=FALSE, fig.keep='all', fig.width=10, fig.height=6}
for (b in levels(stats$Benchmark)) {
  cat(b)
  data_b <- droplevels(subset(stats, Benchmark == b))
  
  p <- ggplot(data_b, aes(x = VM, y = Time.ms, fill = VM)) +
    geom_bar(stat = "identity") +
    geom_errorbar(aes(ymax = Time.ms + sd, ymin = Time.ms - sd), width=0.25) +
    coord_flip() + theme_bw() + # scale_fill_manual(values=col) +
    theme(legend.position="none") + ggtitle(b)
  print(p)
}
```
