# Performance Overview

## Comparing Peak Performance of the different Language Implementations

### Setup

Basic R setup script.

```{r prepare-data, echo=TRUE, message=TRUE, warning=TRUE, dev='svg'}
# load libraries, the data, and prepare it
if (Sys.getenv("RSTUDIO") == "1") { setwd("/Users/smarr/Projects/PostDoc/FASTXX/are-we-fast-yet/report") }

source("scripts/libraries.R", chdir=TRUE)
data <- load_data_file("data/benchmark.data")
data <- subset(data, select = c(Value, Unit, Benchmark, VM, Iteration))
data <- prepare_vm_names(data)

summary(data)
levels(data$VM)

jit <- droplevels(subset(data, Iteration >= 700 & Iteration <= 999))
mri <- droplevels(subset(data, VM == "MRI22"))

peak <- rbind(jit, mri)

norm <- ddply(peak, ~ Benchmark, transform,
              RuntimeRatio = Value / mean(Value[VM == "GraalEnterprise"]))
stats <- ddply(norm, ~ VM + Benchmark, summarise,
               Time.ms = mean(Value),
               sd      = sd(Value),
               RuntimeFactor = geometric.mean(RuntimeRatio),
               RR.sd         = sd(RuntimeRatio))


```

# Plots

## Runtime Factor over GraalEnterprise (Java benchmark version)

 - GraalEnterprise: top-tier JIT = Enterprise Graal
 - GraalC2: top-tier JIT = C2
 - GraalBasic: top-tier JIT = Basic Graal
 - GraalJS: on Enterprise Graal
 - JRubyTruffle: on Basic Graal
 - JRubyTruffleEnterprise: on Enterprise Graal
 - Node: node.js 5.2
 - SOMns: on Basic Graal

```{r fast-vms-factor, echo=FALSE, dev='png', fig.keep='all', fig.width=10, fig.height=6}
# b <- "Mandelbrot"

fast_stats <- droplevels(subset(stats, VM != "MRI22" & VM != "JRubyGraal" & VM != "JRubyC2" & Benchmark != "DeltaBlueV8"))

for (b in levels(fast_stats$Benchmark)) {
  data_b <- droplevels(subset(fast_stats, Benchmark == b))
  
  p <- ggplot(data_b, aes(x = VM, y = RuntimeFactor, fill = VM)) +
    geom_bar(stat = "identity") +
    geom_errorbar(aes(ymax = RuntimeFactor + RR.sd, ymin = RuntimeFactor - RR.sd), width=0.25) +
    coord_flip() + theme_bw() + # scale_fill_manual(values=col) +
    theme(legend.position="none") + ggtitle(b)
  print(p)
  
}
```

## Same Plots: Absolute Runtime

```{r fast-vms-ms, echo=FALSE, dev='png', fig.keep='all', fig.width=10, fig.height=6}
for (b in levels(fast_stats$Benchmark)) {
  data_b <- droplevels(subset(fast_stats, Benchmark == b))
  
  p <- ggplot(data_b, aes(x = VM, y = Time.ms, fill = VM)) +
    geom_bar(stat = "identity") +
    geom_errorbar(aes(ymax = Time.ms + sd, ymin = Time.ms - sd), width=0.25) +
    coord_flip() + theme_bw() + # scale_fill_manual(values=col) +
    theme(legend.position="none") + ggtitle(b)
  print(p)
}
```

# Including Slow Implementations

## Runtime Factor over GraalEnterprise (Java benchmark version)

```{r all-vms-factor, echo=FALSE, dev='png', fig.keep='all', fig.width=10, fig.height=6}
# b <- "Mandelbrot"

levels(stats$Benchmark)

for (b in levels(stats$Benchmark)) {
  cat(b)
  data_b <- droplevels(subset(stats, Benchmark == b))
  
  p <- ggplot(data_b, aes(x = VM, y = RuntimeFactor, fill = VM)) +
    geom_bar(stat = "identity") +
    geom_errorbar(aes(ymax = RuntimeFactor + RR.sd, ymin = RuntimeFactor - RR.sd), width=0.25) +
    coord_flip() + theme_bw() + # scale_fill_manual(values=col) +
    theme(legend.position="none") + ggtitle(b)
  print(p)
  
}
```

## Same Plots: Absolute Runtime

```{r all-vms, echo=FALSE, dev='png', fig.keep='all', fig.width=10, fig.height=6}
for (b in levels(stats$Benchmark)) {
  cat(b)
  data_b <- droplevels(subset(stats, Benchmark == b))
  
  p <- ggplot(data_b, aes(x = VM, y = Time.ms, fill = VM)) +
    geom_bar(stat = "identity") +
    geom_errorbar(aes(ymax = Time.ms + sd, ymin = Time.ms - sd), width=0.25) +
    coord_flip() + theme_bw() + # scale_fill_manual(values=col) +
    theme(legend.position="none") + ggtitle(b)
  print(p)
}
```
