# Performance Overview

## Comparing Peak Performance of Language Implementations

```{r prepare-data, echo=FALSE, message=TRUE, warning=TRUE, dev='svg'}
## Basic Setup
# load libraries, the data, and prepare it
if (Sys.getenv("RSTUDIO") == "1") { setwd("/Users/smarr/Projects/PostDoc/FASTXX/are-we-fast-yet/report") }

source("scripts/libraries.R", chdir=TRUE)
data <- load_data_file("data/benchmark.data")
data <- subset(data, select = c(Value, Unit, Benchmark, VM, Iteration))
data <- prepare_vm_names(data)

# summary(data)
# levels(data$VM)

jit <- droplevels(subset(data, Iteration >= 700 & Iteration <= 999))
mri <- droplevels(subset(data, VM == "MRI22"))

peak <- rbind(jit, mri)

norm <- ddply(peak, ~ Benchmark, transform,
              RuntimeRatio = Value / mean(Value[VM == "GraalEnterprise"]))
stats <- ddply(norm, ~ VM + Benchmark, summarise,
               Time.ms = mean(Value),
               sd      = sd(Value),
               RuntimeFactor = geometric.mean(RuntimeRatio),
               RR.sd         = sd(RuntimeRatio))
stats_all <- stats
stats <- droplevels(subset(stats, Benchmark != "DeltaBlueV8"))

plot_benchmarks_speedup_for_vms <- function(stats, vms) {
  vm_stats <- droplevels(subset(stats, VM %in% vms))
  
  for (b in levels(vm_stats$Benchmark)) {
    data_b <- droplevels(subset(vm_stats, Benchmark == b))
    
    p <- ggplot(data_b, aes(x = VM, y = RuntimeFactor, fill = VM)) +
      geom_bar(stat = "identity") +
      geom_errorbar(aes(ymax = RuntimeFactor + RR.sd, ymin = RuntimeFactor - RR.sd), width=0.25) +
      coord_flip() + theme_bw() + # scale_fill_manual(values=col) +
      theme(legend.position="none") + ggtitle(b)
    tryCatch({print(p)})
  }
}
```

The following plots show the runtime factor over GraalEnterprise (Java benchmark version)

 - GraalEnterprise: top-tier JIT = Enterprise Graal
 - GraalC2: top-tier JIT = C2
 - GraalBasic: top-tier JIT = Basic Graal
 - GraalJS: on Enterprise Graal
 - JRubyTruffle: on Basic Graal
 - JRubyTruffleEnterprise: on Enterprise Graal
 - JRubyTruffleMaster07: JRuby master branch on Graal 0.7 release
 - Node: node.js 5.2
 - SOMns: on Basic Graal

The benchmarks are implemented as identical as possible, trying to exercise a
'core' language common to the benchmarked system.
More details on the methodology and all benchmarks are available in the
[Are We Fast Yet?](https://github.com/smarr/are-we-fast-yet#are-we-fast-yet) repository.

### Are Truffle Languages Fast Yet? (Compared to Java with Graal Enterprise)

```{r truffle-langs, echo=FALSE, dev='png', fig.keep='all', fig.width=10, fig.height=6}
plot_benchmarks_speedup_for_vms(stats, c("GraalEnterprise", "GraalJS", "JRubyTruffle", "JRubyTruffleEnterprise", "JRubyTruffleMaster07", "SOMns"))
```

### Is Graal Fast Yet? (Compared to C2)

```{r graal, echo=FALSE, dev='png', fig.keep='all', fig.width=10, fig.height=6}
plot_benchmarks_speedup_for_vms(stats, c("GraalBasic", "GraalC2", "GraalEnterprise"))
```

### Is Graal Fast Yet? (JRuby-Indy on C2 or Graal)

```{r graal-jruby-indy, echo=FALSE, dev='png', fig.keep='all', fig.width=10, fig.height=6}
plot_benchmarks_speedup_for_vms(stats, c("JRubyC2", "JRubyGraal"))
```

### Is Graal.JS Fast Yet?
```{r graaljs, echo=FALSE, dev='png', fig.keep='all', fig.width=10, fig.height=6}
plot_benchmarks_speedup_for_vms(stats, c("GraalEnterprise", "GraalJS", "Node"))
```

### Is JRuby+Truffle Fast Yet?

```{r jruby-truffle, echo=FALSE, dev='png', fig.keep='all', fig.width=10, fig.height=6}
plot_benchmarks_speedup_for_vms(stats, c("GraalEnterprise", "JRubyTruffle", "JRubyTruffleEnterprise", "JRubyTruffleMaster07"))
```

Including JRuby-Indy (w/o Truffle) and MRI
```{r jruby-truffle-all, echo=FALSE, dev='png', fig.keep='all', fig.width=10, fig.height=6}
plot_benchmarks_speedup_for_vms(stats, c("GraalEnterprise", "JRubyC2", "JRubyGraal", "JRubyTruffle", "JRubyTruffleEnterprise", "JRubyTruffleMaster07", "MRI22"))
```




# Other Plots

## Absolute Runtime

```{r fast-vms-ms, echo=FALSE, dev='png', fig.keep='all', fig.width=10, fig.height=6}
for (b in levels(fast_stats$Benchmark)) {
  data_b <- droplevels(subset(fast_stats, Benchmark == b))
  
  p <- ggplot(data_b, aes(x = VM, y = Time.ms, fill = VM)) +
    geom_bar(stat = "identity") +
    geom_errorbar(aes(ymax = Time.ms + sd, ymin = Time.ms - sd), width=0.25) +
    coord_flip() + theme_bw() + # scale_fill_manual(values=col) +
    theme(legend.position="none") + ggtitle(b)
  print(p)
}
```

## Including Slow Implementations

```{r all-vms, echo=FALSE, dev='png', fig.keep='all', fig.width=10, fig.height=6}
for (b in levels(stats$Benchmark)) {
  cat(b)
  data_b <- droplevels(subset(stats, Benchmark == b))
  
  p <- ggplot(data_b, aes(x = VM, y = Time.ms, fill = VM)) +
    geom_bar(stat = "identity") +
    geom_errorbar(aes(ymax = Time.ms + sd, ymin = Time.ms - sd), width=0.25) +
    coord_flip() + theme_bw() + # scale_fill_manual(values=col) +
    theme(legend.position="none") + ggtitle(b)
  print(p)
}
```
