# Performance Results

This report was generated on `r Sys.time()`.

```{r load-scripts, echo=FALSE, include=FALSE}
# load libraries, the data, and prepare it
if (Sys.getenv("USER") == "smarr") {
  setwd("/Users/smarr/Projects/ast-vs-bc-interp/are-we-fast-yet/report/bc-vs-ast")
  } else if (Sys.getenv("USER") == "sopi") {
  setwd("/home/sopi/Documents/Writing/2023/ast-vs-bc-interp/are-we-fast-yet/report/bc-vs-ast/" )
  } else {
  setwd("/home/octavel/PhD/papers/ast-vs-bc-interp/are-we-fast-yet/report/bc-vs-ast")
  }
source("scripts/libraries.R", chdir=TRUE)

# data <- rbind(
#   # main run, one invocation of everything
#   load_data_url("https://rebench.stefan-marr.de/rebenchdb/get-exp-data/1518"),
#   # another run, adding JDK8
#   load_data_url("https://rebench.stefan-marr.de/rebenchdb/get-exp-data/1523"))
# 
# data <- rbind(
#   data %>% filter(exe != "TruffleSOM-ast-NativeEE-int-uber"),
#   # latest uber run
#   load_data_url("https://rebench.stefan-marr.de/rebenchdb/get-exp-data/1543")) %>%
#   factorize_result()

# data <- 
#   rbind(load_data_url("https://rebench.stefan-marr.de/rebenchdb/get-exp-data/1806") %>%
#           factorize_result(),
#         # PyPy-int
#         load_data_url("https://rebench.stefan-marr.de/rebenchdb/get-exp-data/1814"))


supernodes_data <- load_data_url("https://rebench.stefan-marr.de/TruffleSOM/data/2564") %>%
  mutate(exe = paste0(exe, "-supernodes")) %>%
  mutate(extraargs = ifelse(extraargs == "None", 42, extraargs)) %>%
  filter(criterion == "total")

jdk20_data <- load_data_url("https://rebench.stefan-marr.de/Are-We-Fast-Yet/data/2800")
jdk20_data$exe <- ifelse(grepl("Java-int", jdk20_data$exe), "Java20-int", "Java20-C2-jit")

data <- rbind(
  load_data_url("https://rebench.stefan-marr.de/rebenchdb/get-exp-data/2257") |>
    filter(exe != "CPython-int"),
  load_data_url("https://rebench.stefan-marr.de/rebenchdb/get-exp-data/2258"),
  load_data_url("https://rebench.stefan-marr.de/rebenchdb/get-exp-data/2259"),
  load_data_url("https://rebench.stefan-marr.de/rebenchdb/get-exp-data/2260"),
  load_data_url("https://rebench.stefan-marr.de/rebenchdb/get-exp-data/2275"),
  load_data_url("https://rebench.stefan-marr.de/rebenchdb/get-exp-data/2277"),
  load_data_url("https://rebench.stefan-marr.de/rebenchdb/get-exp-data/2280"),
  load_data_url("https://rebench.dev/Are-We-Fast-Yet/data/2545"),
  jdk20_data,
  supernodes_data
  ) |>
  factorize_result()

data <- rbind(
  load_data_url("https://rebench.stefan-marr.de/Are-We-Fast-Yet/data/2858"),
  jdk20_data
)

# ragg_png = function(..., res = 192) {
#   ragg::agg_png(..., res = res, units = "in")
# }
opts_chunk$set(dev = 'png',
               #fig.ext = 'png',
               #dpi = 120,
               fig.retina = 2,
               dev.args=list(pointsize=10),
               echo = FALSE,
               fig.keep='all',
               fig.path="figures/",
               external=FALSE,
               tidy=FALSE)
#    cache=TRUE,

# dput(levels(data$bench))
vms_all <- names(vm_names)
vms_int <- c(
  "Java-int",
  "Java20-int",
  "Java8-int",
  "Node-int",
  "GraalJS-NativeEE-int",
  
  "CPython-int",
  "GraalPython-NativeEE-int", 
  "PyPy-int",
  
  "CRuby-int",
  "TruffleRuby-NativeEE-int",
 
  "PySOM-ast-int",
  "PySOM-bc-int", 
  
  "TruffleSOM-ast-NativeEE-int-main",
  "TruffleSOM-ast-NativeEE-int-super", 
  "TruffleSOM-ast-NativeEE-int-uber", 
  "TruffleSOM-bc-NativeEE-int-main",
  
  "TruffleSOM-ast-NativeCE-int-astvsbc",
  "TruffleSOM-bc-NativeCE-int-astvsbc",
  
  "SOMpp-int",
  "CSOM-int",
  "SOM-RS-ast-int",
  "SOM-RS-bc-int",
  "ykSOM-int")

vms_jit <- c(
  "Java20-C2-jit",
  "Java17-C2-jit",
  "Java8-C2-jit",
  "Java-native",
  "CSharp-jit",
  "Node-jit",
  "GraalJS-HotspotEE-jit",
  
  "PyPy-jit",
  "GraalPython-HotspotEE-jit", 
  
  "CRuby-y-jit",
  "TruffleRuby-HotspotEE-jit",
  
  "PySOM-ast-jit",
  "PySOM-bc-jit", 
  "TruffleSOM-ast-HotspotCE-jit-main", 
  "TruffleSOM-bc-HotspotCE-jit-main",
  "TruffleSOM-ast-HotspotEE-jit-main", 
  "TruffleSOM-bc-HotspotEE-jit-main"
)

assert_that(all(sort(c(vms_int, vms_jit)) == sort(vms_all))) ## sanity check

warmup_jit_iteration <- 500
baseline_vm_peak <- "Java20-C2-jit"
baseline_vm_int <- "Java20-int"
baseline_vm_first <- "Java20-C2-jit"

results_peak <- data %>%
  compute_all(
    iteration >= warmup_jit_iteration | exe %in% vms_int | exe == "CRuby-y-jit",
    baseline_vm_peak)

results_int <- data %>%
  compute_all(
    exe %in% vms_int,
    baseline_vm_int)

plot_benchmarks_speedup_for_vms <- function(
  vm_norm, vms, label = "Runtime Factor, normalized to Java\n(lower is better)") {
  # vm_norm <- norm_peak
  # vms <- vms_jit
  
  vm_norm <- vm_norm %>%
    filter(exe %in% vms) %>%
    droplevels()
  suppressMessages(vm_norm$exe <- revalue(vm_norm$exe, vm_names) %>% droplevels())
  # TODO: is norm_peak the right table? we don't have exe_ratio there yet
  # vm_norm$exe <- reorder(vm_norm$exe, X=vm_norm$exe_ratio)

  breaks <- levels(droplevels(vm_norm)$exe)
  col_values <- sapply(breaks, function(x) color_set_vms[[x]])
  
  for (b in levels(vm_norm$bench)) {
    # b <- "Bounce"
    data_b <- droplevels(filter(vm_norm, bench == b))

    p <- ggplot(data_b, aes(y = exe, x = ratio_median, fill = exe)) +
      geom_vline(aes(xintercept=1), colour="#333333", linetype="solid") +
      geom_boxplot(aes(color = exe, fill = exe),
                    alpha=0.4,
                   outlier.size = 0.9,
                   outlier.alpha = 0.6,
                   lwd=0.2) +
      geom_jitter(aes(color = exe, y = exe), size=0.3, alpha=0.3) +
      scale_x_log10() +
      scale_y_discrete(limits = rev) +
      scale_color_manual(values = color_set_vms) +
      scale_fill_manual(values = color_set_vms) +
      theme_simple() + # scale_fill_manual(values=col) +
      theme(legend.position="none",
            axis.title.x = element_text(
              size = 12,
              margin = margin(t = 0.1, unit = "cm"))) +
      ggtitle(b) +
      ylab(NULL) +
      xlab(label) 
      
    tryCatch({print(p)})
  }
}

overview_box_plot <- function(stats, vms = NULL, prepare_data = NULL, pre_plot = NULL, new_colors = FALSE, x_breaks = waiver()) {
  # stats <- stats_latest
  # vms <- c("Node", "Pharo", "JavaInt", 
  #          "Lua53", "LuaJIT2", "SOMns-Enterprise")

  if (is.null(vms)) {
    vm_stats <- stats
  } else {
    vm_stats <- stats %>%
      filter(exe %in% vms) %>%
      droplevels()
  }

  if (!is.null(prepare_data)) {
   vm_stats <- vm_stats %>%
     prepare_data() %>%
     droplevels()
  }

  suppressMessages(vm_stats$exe <- revalue(vm_stats$exe, vm_names))
  breaks <- levels(vm_stats$exe)

  if (new_colors) {
    col_values <- brewer.pal(length(breaks), "Paired")
  } else {
    col_values <- sapply(breaks, function(x) color_set_vms[[x]])
  }

  median_fn <- median
  
#col_names <- names(vm_stats)

# Loop over each column in the data
#for (col in col_names) {
  # Check for missing values (NA) in each column
#  if (any(is.na(vm_stats[[col]]))) {
#    cat(paste("Column", col, "has missing values\n"))
#  }
  # Check for non-finite values (Inf, -Inf, NaN) in each column
#  if (!all(is.finite(vm_stats[[col]]))) {
#    cat(paste("Column", col, "has non-finite values\n"))
#  }
#}
  
  # Only two decimal places
  x_breaks <- round(x_breaks, 2)
  
  plot <- ggplot(vm_stats, aes(x=ratio, y=reorder(exe, -ratio, FUN = median_fn)))
  if (!is.null(pre_plot)) {
    plot <- pre_plot(plot)
  }
  
  if (!identical(x_breaks, waiver())) {
    for (b in x_breaks) {
      # REM: the loop requires the use of the eager aes_ here!
      plot <- plot + geom_vline(aes(xintercept=b),
                                colour="#cccccc", linetype="dashed")
    }
  }
  plot <- plot + 
    geom_vline(aes(xintercept=1),    colour="#999999", linetype="solid") +
    geom_vline(aes(xintercept=2), colour="#999999", linetype="dotted") +
    geom_vline(aes(xintercept=10), colour="#999999", linetype="dotted")
    
  plot <- plot +
    geom_boxplot(outlier.size = 0.5,
                 aes(colour = exe, fill = exe), 
                  alpha=0.4) + #fill=get_color(5, 7)
    scale_x_log10(breaks = x_breaks) +
    scale_y_discrete(label = function (x) parse(text=x)) +
    scale_color_manual(values = color_set_vms) +
    scale_fill_manual(values = color_set_vms) +
    theme_bw() + theme_simple(font_size = 8) +
    theme(
      axis.text.x = element_text(angle= 90, vjust=0.5, hjust=1),
      axis.title.x  = element_text(size = 8, family="Arial"),
      legend.position="none") +
    #scale_y_log10(breaks=c(1,2,3,10,20,30,50,100,200,300,500,1000)) + #limit=c(0,30), breaks=seq(0,100,5), expand = c(0,0)
    ylab("") +
    xlab("")
    #xlab("Runtime Factor, normalized to Java (lower is better)")
  plot
}

```

For VMs with just-in-time compilation, we report results based on
the measurements starting from iteration `r warmup_jit_iteration`.
While this does not guarantee that we measure "peak performance",
it gives us relatively stable results for most language implementations.

## Overview

###### TruffleSOM/PySOM JIT Peak Performance

Results are normalized either to `r vm_names[baseline_vm_peak]`.

```{r tsom-pysom-jit, fig.width=5.5, fig.height=1.2}
plot <- overview_box_plot(results_peak$stats$bench,
                  c("PySOM-ast-jit", "PySOM-bc-jit", 
                    "TruffleSOM-ast-HotspotCE-jit-main",
                    "TruffleSOM-bc-HotspotCE-jit-main", "Java20-C2-jit", "Node-jit"),
                  x_breaks = c(0.5, 0.75, 1, 1.33333, 2, 3, 5, 10, 20))
ggsave("../../../paper/images/tsom-pysom-peak-perf.pdf", plot, device= cairo_pdf)
plot
```

###### TruffleSOM/PySOM Interpreter Performance

Comparing TruffleSOM, PySOM, Node-int and Java-Int. Using `r vm_names[baseline_vm_int]` as a baseline.

```{r tsom-pysom-int, fig.width=5.5, fig.height=1.2, cache=FALSE}
plot <- overview_box_plot(results_int$stats$bench,
                  c("PySOM-ast-int", "PySOM-bc-int",
                    "TruffleSOM-ast-NativeCE-int-astvsbc", "TruffleSOM-bc-NativeCE-int-astvsbc",
                    "TruffleSOM-native-interp-bc-supernodes",
                    "Java20-int", "Node-int"),
                  x_breaks = c(0.41, 0.5, 0.75, 1, 1.33333, 2, 3, 5, 10))

ggsave("../../../paper/images/tsom-pysom-interp-perf.pdf", plot, device=cairo_pdf)
plot
```


## Write to text file

```{r write-file}
X2 <- function(num) {
  format(num, digits=2, nsmall = 2)
}

WithT <- function(num) {
  format(num, big.mark = ",")
}

tex.file <- file("../../../paper/gen/overview.tex")

# Numbers for part 5.1
min_row = results_int$stats$exe[which.min(results_int$stats$exe$min_ratio),]
min_exe <- min_row$exe
speedup <- (min_row$min_ratio - 1) * 100

macros = c(
  paste0("\\newcommand{\\LowestRatioIntJava}{",
        X2(abs(speedup)), "\\xspace}"),
  paste0("\\newcommand{\\SpeediestExeInt}{",
        X2(min_exe), "\\xspace}"),
  paste0("\\newcommand{\\PySOMastMedian}{",
        X2((results_int$stats$exe |> filter(exe == "PySOM-ast-int"))$median_ratio), "\\xspace}"),
  paste0("\\newcommand{\\PySOMbcMedian}{",
        X2((results_int$stats$exe |> filter(exe == "PySOM-bc-int"))$median_ratio), "\\xspace}"),
  paste0("\\newcommand{\\slowestPySOMast}{",
        X2((results_int$stats$exe |> filter(exe == "PySOM-ast-int"))$max_ratio), "\\xspace}"),
  paste0("\\newcommand{\\slowestPySOMbc}{",
        X2((results_int$stats$exe |> filter(exe == "PySOM-bc-int"))$max_ratio), "\\xspace}"),
  paste0("\\newcommand{\\fastestTSOMast}{",
        X2((results_int$stats$exe |> filter(exe == "TruffleSOM-ast-NativeCE-int-astvsbc"))$min_ratio), "\\xspace}"),
  paste0("\\newcommand{\\fastestTSOMbc}{",
        X2((results_int$stats$exe |> filter(exe == "TruffleSOM-bc-NativeCE-int-astvsbc"))$min_ratio), "\\xspace}"),
  paste0("\\newcommand{\\TSOMastMedian}{",
        X2((results_int$stats$exe |> filter(exe == "TruffleSOM-ast-NativeCE-int-astvsbc"))$median_ratio), "\\xspace}"),
  paste0("\\newcommand{\\TSOMbcMedian}{",
        X2((results_int$stats$exe |> filter(exe == "TruffleSOM-bc-NativeCE-int-astvsbc"))$median_ratio), "\\xspace}")
)

# Numbers for part 5.2
only_soms_jit <- results_peak$stats$exe |> filter(grepl('SOM', exe), grepl('jit', exe)) |> arrange(median_ratio)
min_row_jit = only_soms_jit[which.min(only_soms_jit$median_ratio),]
max_row_jit = only_soms_jit[which.max(only_soms_jit$median_ratio),]

macros = c(macros,
  paste0("\\newcommand{\\SpeediestSOMjitMedianFirst}{",
        X2(min_row_jit$median_ratio), "\\xspace}"),
  paste0("\\newcommand{\\SpeediestSOMjitExeFirst}{",
        X2(min_row_jit$exe), "\\xspace}"),
  paste0("\\newcommand{\\SpeediestSOMjitMedianSecond}{",
        X2(only_soms_jit[2,]$median_ratio), "\\xspace}"),
  paste0("\\newcommand{\\SpeediestSOMjitExeSecond}{",
        X2(only_soms_jit[2,]$exe), "\\xspace}"),
  paste0("\\newcommand{\\SlowliestSOMjitMedianThird}{",
        X2(only_soms_jit[3,]$median_ratio), "\\xspace}"),
  paste0("\\newcommand{\\SlowliestSOMjitExeThird}{",
        X2(only_soms_jit[3,]$exe), "\\xspace}"),
  paste0("\\newcommand{\\SlowliestSOMjitMedianFourth}{",
        X2(max_row_jit$median_ratio), "\\xspace}"),
  paste0("\\newcommand{\\SlowliestSOMjitExeFourth}{",
        X2(max_row_jit$exe), "\\xspace}"),
  paste0("\\newcommand{\\PySOMastLowerBound}{",
        X2((only_soms_jit |> filter(exe == "PySOM-ast-jit"))$min_ratio), "\\xspace}"),
  paste0("\\newcommand{\\PySOMastUpperBound}{",
        X2((only_soms_jit |> filter(exe == "PySOM-ast-jit"))$max_ratio), "\\xspace}"),
  paste0("\\newcommand{\\PySOMbcLowerBound}{",
        X2((only_soms_jit |> filter(exe == "PySOM-bc-jit"))$min_ratio), "\\xspace}"),
  paste0("\\newcommand{\\PySOMbcUpperBound}{",
        X2((only_soms_jit |> filter(exe == "PySOM-bc-jit"))$max_ratio), "\\xspace}"),
  paste0("\\newcommand{\\TSOMastLowerBound}{",
        X2((only_soms_jit |> filter(exe == "TruffleSOM-ast-HotspotCE-jit-main"))$min_ratio), "\\xspace}"),
  paste0("\\newcommand{\\TSOMastUpperBound}{",
        X2((only_soms_jit |> filter(exe == "TruffleSOM-ast-HotspotCE-jit-main"))$max_ratio), "\\xspace}"),
  paste0("\\newcommand{\\TSOMbcLowerBound}{",
        X2((only_soms_jit |> filter(exe == "TruffleSOM-bc-HotspotCE-jit-main"))$min_ratio), "\\xspace}"),
  paste0("\\newcommand{\\TSOMbcUpperBound}{",
        X2((only_soms_jit |> filter(exe == "TruffleSOM-bc-HotspotCE-jit-main"))$max_ratio), "\\xspace}")
)


writeLines(macros, tex.file)
close(tex.file)
```