# Performance Results

This report was generated on `r Sys.time()`.

```{r load-scripts, echo=FALSE, include=FALSE}
# load libraries, the data, and prepare it
if (Sys.getenv("USER") == "smarr") {
  setwd("/Users/smarr/Projects/ast-vs-bc-interp/are-we-fast-yet/report/bc-vs-ast")
} else if (Sys.getenv("USER") == "sopi") {
  setwd("/home/sopi/Documents/Writing/2023/ast-vs-bc-interp/are-we-fast-yet/report/bc-vs-ast/" )
} else {
  setwd("/home/octavel/PhD/papers/ast-vs-bc-interp/are-we-fast-yet/report/bc-vs-ast")
}
source("scripts/libraries.R", chdir=TRUE)

jit_data <- load_data_url("https://rebench.stefan-marr.de/Are-We-Fast-Yet/data/3048")
jit_data$exe <- as.character(jit_data$exe)
jit_data$exe[jit_data$exe=="TruffleSOM-ast-HotspotCE-jit"] <- "TruffleSOM-ast-HotspotCE-jit-main" # to make it fit the old data
jit_data$exe[jit_data$exe=="TruffleSOM-bc-HotspotCE-jit"] <- "TruffleSOM-bc-HotspotCE-jit-main"


interp_exes <- c("PySOM-bc-int", "Node-int", "PySOM-ast-int", "PyPy-int", "Java-int", "TruffleSOM-bc-NativeCE-int-astvsbc", "TruffleSOM-ast-NativeCE-int-astvsbc")
interp_data <- load_data_url("https://rebench.stefan-marr.de/Are-We-Fast-Yet/data/2987")
interp_data$exe <- as.character(interp_data$exe)
interp_data$exe[interp_data$exe=="TruffleSOM-ast-NativeCE-int"] <- "TruffleSOM-ast-NativeCE-int-astvsbc" # to make it fit the old data
interp_data$exe[interp_data$exe=="TruffleSOM-bc-NativeCE-int"] <- "TruffleSOM-bc-NativeCE-int-astvsbc"

data <- rbind(jit_data, interp_data)

# old
#data <- rbind(data %>% filter(!(exe %in% interp_exes)), interp_data)

# ragg_png = function(..., res = 192) {
#   ragg::agg_png(..., res = res, units = "in")
# }
opts_chunk$set(dev = 'png',
               #fig.ext = 'png',
               #dpi = 120,
               fig.retina = 2,
               dev.args=list(pointsize=10),
               echo = FALSE,
               fig.keep='all',
               fig.path="figures/",
               external=FALSE,
               tidy=FALSE)
#    cache=TRUE,

# dput(levels(data$bench))
vms_all <- names(vm_names)
vms_int <- c(
  "Java-int",
  "Java20-int",
  "Java8-int",
  "Node-int",
  "GraalJS-NativeEE-int",
  
  "CPython-int",
  "GraalPython-NativeEE-int", 
  "PyPy-int",
  
  "CRuby-int",
  "TruffleRuby-NativeEE-int",
 
  "PySOM-ast-int",
  "PySOM-bc-int", 
  
  "TruffleSOM-ast-NativeEE-int-main",
  "TruffleSOM-ast-NativeEE-int-super", 
  "TruffleSOM-ast-NativeEE-int-uber", 
  "TruffleSOM-bc-NativeEE-int-main",
  
  "TruffleSOM-ast-NativeCE-int-astvsbc",
  "TruffleSOM-bc-NativeCE-int-astvsbc",
  
  "SOMpp-int",
  "CSOM-int",
  "SOM-RS-ast-int",
  "SOM-RS-bc-int",
  "ykSOM-int")

warmup_jit_iteration <- 1000
baseline_vm_peak <- "Java20-C2-jit"
baseline_vm_int <- "Java20-int"
baseline_vm_first <- "Java20-C2-jit"

results_peak <- data %>%
  compute_all(
    iteration >= warmup_jit_iteration | exe %in% vms_int,
    baseline_vm_peak)

results_int <- data %>%
  compute_all(
    exe %in% vms_int,
    baseline_vm_int)

plot_benchmarks_speedup_for_vms <- function(
  vm_norm, vms, label = "Runtime Factor, normalized to Java\n(lower is better)") {
  # vm_norm <- norm_peak
  # vms <- vms_jit
  
  vm_norm <- vm_norm %>%
    filter(exe %in% vms) %>%
    droplevels()
  suppressMessages(vm_norm$exe <- revalue(vm_norm$exe, vm_names) %>% droplevels())
  # TODO: is norm_peak the right table? we don't have exe_ratio there yet
  # vm_norm$exe <- reorder(vm_norm$exe, X=vm_norm$exe_ratio)

  breaks <- levels(droplevels(vm_norm)$exe)
  col_values <- sapply(breaks, function(x) color_set_vms[[x]])
  
  for (b in levels(vm_norm$bench)) {
    # b <- "Bounce"
    data_b <- droplevels(filter(vm_norm, bench == b))

    p <- ggplot(data_b, aes(y = exe, x = ratio_median, fill = exe)) +
      geom_vline(aes(xintercept=1), colour="#333333", linetype="solid") +
      geom_boxplot(aes(color = exe, fill = exe),
                    alpha=0.4,
                   outlier.size = 0.9,
                   outlier.alpha = 0.6,
                   lwd=0.2) +
      geom_jitter(aes(color = exe, y = exe), size=0.3, alpha=0.3) +
      scale_x_log10() +
      scale_y_discrete(limits = rev) +
      scale_color_manual(values = color_set_vms) +
      scale_fill_manual(values = color_set_vms) +
      theme_simple() + # scale_fill_manual(values=col) +
      theme(legend.position="none",
            axis.title.x = element_text(
              size = theme_simple_font_size(),
              margin = margin(t = 0.1, unit = "cm"))) +
      ggtitle(b) +
      ylab(NULL) +
      xlab(label) 
      
    tryCatch({print(p)})
  }
}

overview_box_plot <- function(stats, vms = NULL, prepare_data = NULL, pre_plot = NULL, new_colors = FALSE, x_breaks = waiver(), x_lab = NULL, y_lab = NULL) {
  # stats <- stats_latest
  # vms <- c("Node", "Pharo", "JavaInt", 
  #          "Lua53", "LuaJIT2", "SOMns-Enterprise")

  if (is.null(vms)) {
    vm_stats <- stats
  } else {
    vm_stats <- stats %>%
      filter(exe %in% vms) %>%
      droplevels()
  }

  if (!is.null(prepare_data)) {
   vm_stats <- vm_stats %>%
     prepare_data() %>%
     droplevels()
  }

  suppressMessages(vm_stats$exe <- revalue(vm_stats$exe, vm_names))
  breaks <- levels(vm_stats$exe)

  # if (new_colors) {
  #   col_values <- brewer.pal(length(breaks), "Paired")
  # } else {
  #   col_values <- sapply(breaks, function(x) color_set_vms[[x]])
  # }

  median_fn <- median
  
#col_names <- names(vm_stats)

# Loop over each column in the data
#for (col in col_names) {
  # Check for missing values (NA) in each column
#  if (any(is.na(vm_stats[[col]]))) {
#    cat(paste("Column", col, "has missing values\n"))
#  }
  # Check for non-finite values (Inf, -Inf, NaN) in each column
#  if (!all(is.finite(vm_stats[[col]]))) {
#    cat(paste("Column", col, "has non-finite values\n"))
#  }
#}
  
  # Only two decimal places
  x_breaks <- round(x_breaks, 2)
  
  plot <- ggplot(vm_stats, aes(x=ratio, y=reorder(exe, -ratio, FUN = median_fn)))
  if (!is.null(pre_plot)) {
    plot <- pre_plot(plot)
  }
  
  if (!identical(x_breaks, waiver())) {
    for (b in x_breaks) {
      # REM: the loop requires the use of the eager aes_ here!
      plot <- plot + geom_vline(aes(xintercept=b),
                                colour="#cccccc", linetype="dashed")
    }
  }
  plot <- plot + 
    geom_vline(aes(xintercept=1),    colour="#999999", linetype="solid") +
    geom_vline(aes(xintercept=2), colour="#999999", linetype="dotted") +
    geom_vline(aes(xintercept=10), colour="#999999", linetype="dotted")
    
  plot <- plot +
    geom_boxplot(outlier.size = 0.5,
                 aes(colour = exe, fill = exe), 
                  alpha=0.4) + #fill=get_color(5, 7)
    scale_x_log10(breaks = x_breaks) +
    scale_y_discrete(label = function (x) parse(text=x)) +
    scale_color_manual(values = color_set_vms) +
    scale_fill_manual(values = color_set_vms) +
    theme_simple() +
    theme(
      axis.text.x = element_text(angle= 90, vjust=0.5, hjust=1),
      axis.title.x = theme_simple_axis_title(),
      legend.position="none")
    #scale_y_log10(breaks=c(1,2,3,10,20,30,50,100,200,300,500,1000)) + #limit=c(0,30), breaks=seq(0,100,5), expand = c(0,0)
    #xlab("Runtime Factor, normalized to Java (lower is better)")
  
  if (is.null(x_lab)) {
    plot <- plot + xlab("")
  } else {
    plot <- plot + xlab(x_lab)
  }
  
  if (is.null(y_lab)) {
    plot <- plot + ylab("")
  } else {
    plot <- plot + ylab(y_lab)
  }
  
  plot
}

```

For VMs with just-in-time compilation, we report results based on
the measurements starting from iteration `r warmup_jit_iteration`.
While this does not guarantee that we measure "peak performance",
it gives us relatively stable results for most language implementations.

## Overview

###### TruffleSOM/PySOM JIT Peak Performance

Results are normalized either to `r vm_names[baseline_vm_peak]`.

```{r tsom-pysom-jit, fig.width=5.5, fig.height=1.3}
plot <- overview_box_plot(results_peak$stats$bench,
                  c("PySOM-ast-jit", "PySOM-bc-jit", 
                    "TruffleSOM-ast-HotspotCE-jit-main",
                    "TruffleSOM-bc-HotspotCE-jit-main", "Java20-C2-jit", "Node-jit"),
                  x_breaks = c(0.5, 0.75, 1, 1.33333, 2, 3, 5, 10, 20),
                  x_lab = "lower is better")
ggsave("../../../paper/images/tsom-pysom-peak-perf.pdf", plot, device=cairo_pdf)
plot
```

###### TruffleSOM/PySOM Interpreter Performance

Comparing TruffleSOM, PySOM, Node-int and Java-Int. Using `r vm_names[baseline_vm_int]` as a baseline.

```{r tsom-pysom-int, fig.width=5.5, fig.height=1.3, cache=FALSE}
plot <- overview_box_plot(results_int$stats$bench,
                  c("PySOM-ast-int", "PySOM-bc-int",
                    "TruffleSOM-ast-NativeCE-int-astvsbc", "TruffleSOM-bc-NativeCE-int-astvsbc",
                    "Java20-int", "Node-int"),
                  x_breaks = c(0.3, 0.5, 0.75, 1, 1.33333, 2, 3, 5, 10),
                  x_lab = "lower is better")

ggsave("../../../paper/images/tsom-pysom-interp-perf.pdf", plot, device=cairo_pdf)
plot
```


## Write to text file

```{r write-file}
tex.file <- file("../../../paper/gen/overview.tex")

# Numbers for part 5.1
min_row = results_int$stats$exe[which.min(results_int$stats$exe$min_ratio),]
min_exe <- min_row$exe
# speedup <- (min_row$min_ratio - 1) * 100

fastest_bench <- results_int$stats$bench |> filter(exe == min_row$exe, ratio == min_row$min_ratio)


macros = c(
  Cmd("LowestRatioIntJavaPer",       Per(min_row$min_ratio)),
  Cmd("LowestRatioIntJavaBenchmark", fastest_bench$bench),
  
  CmdsMedianMinMaxRatioX("PySOMastInt", results_int$stats$exe |> filter(exe == "PySOM-ast-int")),
  CmdsMedianMinMaxRatioX("PySOMbcInt",  results_int$stats$exe |> filter(exe == "PySOM-bc-int")),

  CmdsMedianMinMaxRatioX("TSOMastInt", results_int$stats$exe |> filter(exe == "TruffleSOM-ast-NativeCE-int-astvsbc")),
  CmdsMedianMinMaxRatioX("TSOMbcInt",  results_int$stats$exe |> filter(exe == "TruffleSOM-bc-NativeCE-int-astvsbc"))
)

# Numbers for part 5.2
only_soms_jit <- results_peak$stats$exe |> filter(grepl('SOM', exe), grepl('jit', exe)) |> arrange(median_ratio)
min_row_jit = only_soms_jit[which.min(only_soms_jit$median_ratio),]
max_row_jit = only_soms_jit[which.max(only_soms_jit$median_ratio),]

macros = c(macros,
  CmdsMedianMinMaxRatioX("NodeJit", results_peak$stats$exe |> filter(exe == "Node-jit")),
  
  CmdsMedianMinMaxRatioX("PySOMastJit", results_peak$stats$exe |> filter(exe == "PySOM-ast-jit")),
  CmdsMedianMinMaxRatioX("PySOMbcJit",  results_peak$stats$exe |> filter(exe == "PySOM-bc-jit")),
  
  CmdsMedianMinMaxRatioX("TSOMastJit", results_peak$stats$exe |> filter(exe == "TruffleSOM-ast-HotspotCE-jit-main")),
  CmdsMedianMinMaxRatioX("TSOMbcJit",  results_peak$stats$exe |> filter(exe == "TruffleSOM-bc-HotspotCE-jit-main"))
)

writeLines(macros, tex.file)
close(tex.file)
```