Performance of SOMns vs Java
============================

```{r echo=FALSE, message=FALSE, warning=FALSE, results='hide', errors=FALSE}
# load libraries, the data, and prepare it
if (Sys.getenv("RSTUDIO") == "1") { setwd("/Users/smarr/Projects/PostDoc/SELF-OPT-INT/performance-overview") }

source("../scripts/libraries.R", chdir=TRUE)
data <- load_data_file("somns-vs-java.data")
data <- subset(data, select = c(Value, Unit, Benchmark, VM, Suite, Extra, Iteration))
data <- prepare_vm_names(data)
# steady <- droplevels(subset(data, grepl("som", Suite) | grepl("java", Suite) | grepl("startup", Suite)))
```

```{r echo=FALSE}
# remove benchmarks that are not availabe for all VMs
data <- subset(data, Benchmark != "TreeSort" & Benchmark != "GraphSearch" & Benchmark != "WhileLoop" & Benchmark != "IntegerLoop" & Benchmark != "FieldLoop")

# aggregate results for display
jit    <- subset(data, Iteration >= 1100 & Iteration <= 1450  & !grepl("Loop", Benchmark))
interp <- subset(data, Iteration >= 25 & grepl("interp", Suite) & !grepl("Loop", Benchmark))

stats_jit <- ddply(jit, ~ Benchmark + VM + Suite,
               summarise,
               Time.mean                 = mean(Value),
               Time.geomean              = geometric.mean(Value),
               Time.stddev               = sd(Value),
               Time.median               = median(Value),
               max = max(Value),
               min = min(Value))
stats_interp <- ddply(interp, ~ Benchmark + VM + Suite,
               summarise,
               Time.mean                 = mean(Value),
               Time.geomean              = geometric.mean(Value),
               Time.stddev               = sd(Value),
               Time.median               = median(Value),
               max = max(Value),
               min = min(Value))

# normalize for each benchmark separately to the Java baseline
norm_jit <- ddply(stats_jit, ~ Benchmark, transform,
              RuntimeRatio = Time.mean / Time.mean[VM == "Java8"])

norm_interp <- ddply(stats_interp, ~ Benchmark, transform,
              RuntimeRatio = Time.mean / Time.mean[VM == "Java8-interp"])

# summarize to VMs
vms_jit <- ddply(norm_jit, ~ VM,
             summarise,
             RunRatio.geomean = geometric.mean(RuntimeRatio),
             min = min(RuntimeRatio),
             max = max(RuntimeRatio),
             type = "Compiled")
vms_interp <- ddply(norm_interp, ~ VM,
             summarise,
             RunRatio.geomean = geometric.mean(RuntimeRatio),
             min = min(RuntimeRatio),
             max = max(RuntimeRatio),
             type = "Interpreted")
```


Overall, SOMns peak performance is about `r round(vms_jit[vms_jit$VM=="SOMns",]$RunRatio.geomean, 2)`x
slower than Java's. The following boxplot summarieses this result.

```{r som-box-overview, echo=FALSE, fig.width=5, fig.height=1.5, dev='svg'}
soms <- droplevels(subset(norm_jit, VM == "SOMns"))


plot <- ggplot(soms, aes(x=VM, y=RuntimeRatio)) +
  geom_hline(aes(yintercept=1), colour="#cccccc", linetype="dashed") +
  geom_hline(aes(yintercept=5), colour="#cccccc", linetype="dashed") +
  geom_boxplot(fill=get_color(5, 7)) + theme_bw() + theme_simple() + theme(axis.text.x = element_text(angle= 90, vjust=0.5, hjust=1)) +
  #scale_y_log10(breaks=c(1,2,3,10,20,30,50,100,200,300,500,1000)) + #limit=c(0,30), breaks=seq(0,100,5), expand = c(0,0)
  ggtitle("Runtime, normalized to Java (lower is better)") + coord_flip() + xlab("")

  #labeller = label_parsed
 plot
```

For this comparison, we use `r length(levels(soms$Benchmark))`
benchmarks. Including classic ones such as DeltaBlue, Richards, Mandelbrot, and
NBody, but also less common ones such as Json, which is a simple JSON parser,
as well as PageRank, which simulates the page-rank algorithm.

```{r compiled, echo=FALSE, fig.width=5, fig.height=3, dev='svg'}
somns <- droplevels(subset(norm_jit, VM == "SOMns"))

p <- ggplot(somns, aes(x=Benchmark, y=RuntimeRatio))
#p <- add_hlines(p, seq(1, 20, 2))
p <- p + geom_boxplot(outlier.size = 0.9) +
  theme_simple() +
  facet_grid(. ~ VM) +
  # facet_grid(. ~ type + VM, labeller = label_parsed) +
  scale_y_continuous(name="Runtime normalized to\nJava (compiled)",
                     breaks=c(1,2,3,4,5))
p <- p + theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust=0.5),
        panel.border = element_rect(colour = "black", fill = NA))
p
```

```{r interpreted, echo=FALSE, fig.width=5, fig.height=3, dev='pdf'}
#p <- ggplot(droplevels(subset(norm_interp, !grepl("Java", VM) & VM != "SOM++-interp")), aes(x=Benchmark, y=RuntimeRatio))
##p <- add_hlines(p, seq(1, 20, 2))
#p <- p + geom_boxplot(outlier.size = 0.9) +
#  theme_simple() +
#  facet_grid(. ~ VM) +
#  # facet_grid(. ~ type + VM, labeller = label_parsed) +
#  scale_y_continuous(name="Runtime normalized to\nJava (interpreted)",
#                     breaks=c(1, 4, 8, 12, 16, 20))
#p <- p + theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust=0.5),
#        panel.border = element_rect(colour = "black", fill = NA))
#p
```




```{r echo=FALSE, fig.width=7, fig.height=5, dev='pdf'}
# create a simple bar chart
# plot <- ggplot(norm, aes_string(x="VM", y="RuntimeRatio"))
# plot <- plot + geom_bar(stat="identity",
#                    colour=get_color(5, 6),
#                    size=.3,        # Thinner lines
#                    fill=get_color(5, 7),
#                    width=0.75) +
#   scale_y_continuous(limit=c(0,30),
#                                                     breaks=seq(0,100,5),
#                                                     expand = c(0,0)) +
#   ylab("Runtime, normalized to Java (lower is better)")
# plot <- plot + facet_wrap(~ Benchmark)
# plot <- plot +
#     theme_bw() + theme_simple() +
#     theme(axis.text.x          = element_text(angle= 90, vjust=0.5, hjust=1))
#   plot

```

```{r echo=FALSE, results='asis'}
table_data <- droplevels(subset(stats_jit, VM == "Java8" | VM == "SOMns"))
name_map <- list("Java8"     = "Java")
levels(table_data$VM)  <- map_names(levels(table_data$VM), name_map)

writeLines("<div class='full center'>")
t <- tabular(Format(digits=2)*Benchmark*Heading()*Time.mean*Justify(data=r)*(Heading()*mean)
             ~ Justify(l,data=r)*Heading('Runtime in ms')*VM,  data=droplevels(table_data))
html(t)
writeLines("</div>")
```

```{r echo=FALSE, results='asis'}
# writeLines("<div class='full center'>")
# t <- tabular(Justify(l,data=l)*VM ~ Format(digits=2)*Heading('Runtime in ms')*Benchmark*Heading()*Value*Justify(data=r)*(mean + sd),  data=droplevels(steady))
# html(t)
# writeLines("</div>")
```

<p>&nbsp</p>
